%************************************************
\chapter{Soldering Experiment}
%************************************************
\begin{flushright}
September 8, 2012
\end{flushright}
\section{Aim}
To learn how to solder and desolder properly, by attaching a 20 pin IC to a test board and removing it after successful soldering.

\section{Theory}
	\subsection {Tools}
		\begin{enumerate}
			\item Solder Station with temperature control
			\item Solder Wire
			\item Wire Stripper
			\item Wire Cutter			
			\item Desoldering Pump
			\item Solder Flux
		\end{enumerate}
	\subsection {Material}
		\begin{enumerate}
			\item IC Holder (20 pin)
			\item Test Board
		\end{enumerate}
\section{`Theory'}
	\begin{enumerate}
		\item Soldering
			\begin{enumerate}
				\item Apply flux to the metal parts intended to be joint
				\item Switch
			\end{enumerate} 
		\item Connected the circuit according to the circuit diagram as given in, keeping the following in mind

			\begin{enumerate}
				\item \emph{Red Wires} are used for the phase connections.
				\item \emph{Black Wires} are used for the neutral connections.
				\item Colours other than Red, Black and Green can be used for the connecting wires.
			\end{enumerate}	
		\item Attached the Lamp Holder using screws \& nut-bolts.
	\end{enumerate}
\section{Precaution}
	\begin{enumerate}
		\item Connections should be tight, viz. shouldn't come out when pulled.
		\item Wires shouldn't be sticking out of the connections without insulation.
		\item Wires should be of appropriate length.
		\item Colours of the wires should be chosen in accordance with their type.
	\end{enumerate}	
\section{Acknowledgements}
I thank Mr. Jatinder Singh for his guidance during the experiment.