\documentclass[11pt]{book}
\renewcommand{\thechapter}{\arabic{chapter}}
%\renewcommand{\chaptername}{}
\newenvironment{project}[3]
{
\chapter{#1}
\noindent{\large \bf Roll Number~:~}{#2}\\
\noindent{\large \bf Name~:~}{#3}
}

\begin{document}
\begin{project}{Quantum Cryptography}{MS11012}{Subhash Gupta}
\section{Introduction and Background}
This section will provide background and information about
the theme. 
Please the format above and give the title, roll number and
your name name in the required format.
The background material is available
in~\cite{ms11012:ref1}. For a general introduction to the
field see~\cite{ms11012:ref2}. A pure state is define as
\begin{equation}
\rho = \vert \psi \rangle \langle \psi \vert
\label{ms11012:eq1}
\end{equation}
As is clear from the definition given in
Equation~(\ref{ms11012:eq1}) we can see that all states are
not pure.
\begin{equation}
\rho = 
\frac{1}{2}\vert \psi \rangle \langle \psi \vert
+\frac{1}{2}\vert \phi \rangle \langle \phi \vert
\label{ms11012:eq2}
\end{equation}
The state given in Equation~(\ref{ms11012:eq2}) is not a pure
state.
\subsection{Remember}
Please remember that all labels for equations will have your
roll number as prefix as given above. This is being done
some that when I put all the latex files together to
generate a book there should be no conflict of labels.
Similarly for references in the bibliography and the
citations labels will have to have your roll number as
prefix. 
\section{This is one of the main section}
You will give this section a name depending upon your
project. The material in section can be divided into
subsection 
\subsection{This is a subsection}
We can have many subsection with headings inside a section
\subsection{This is another one}
\section{This is another main section}
Your project report may require more than one main section!
Feel free to create more section if needed.

\section{Summary and Discussion}
Provide a summary and discuss what you have understood.
summarize the main points and also mention if have found some
subtopics difficult.
\begin{thebibliography}{10}
\bibitem{ms11012:ref1}
Arvind and Arvind, Quantum tomography {\it Physical Review
A}{\bf 73}, 211-215 (2014).
\bibitem{ms11012:ref2}
Neilson and Chuang, Quantum Computation and Quantum
Information ........
\end{thebibliography}
\end{project}
\end{document}